```{r}
library(tidyverse)
library(cfbfastR)
library(Hmisc)
```


```{r}
plays_2023 <- cfbd_pbp_data(2023)
```


```{r}
plays_2023 |> filter(drive_number == 1, play_number == 1, play_type != 'Kickoff') |> distinct(home, play_type)
```
First play: teams are inconsistent about what is defined as a first play. 

```{r}
logs <- read_csv ("https://dwillis.github.io/sports-data-files/footballlogs1122.csv")
```
```{r}
head(logs)
```

```{r}
logs <- logs |> mutate(Differential = TeamScore - OpponentScore)
```


```{r}
logs <- logs %>%
  mutate (
    Differential = abs(TeamScore - OpponentScore)
  )
```

```{r}
correlations <- read_csv("https://dwillis.github.io/sports-data-files/footballlogs1122.csv")
```

```{r}
newcorrelations <- correlations |>
  mutate(
      differential = TeamScore - Penalties
  )
```

```{r}
newcorrelations |> summarise (correlation = cor(differential, Penalties, method="pearson"))
```

```{r}
fit <- lm(differential ~ Penalties, data = newcorrelations)
summary(fit)
```
The p-value is 2.2e-16 so this means the results can be random but are still related in some way. The r-square value is 0.02 which means the differential in score isn't entirely related to how many penalties a team gets. This regression might work better for individual teams (i.e. Maryland who may or may not get bogus calls seemingly every game) to see if penalties impact a teams win/loss record.  


```{r}
simplelogs <- logs |> select_if(is.numeric) |> select(-Game) |> select(Differential, Penalties, TotalTurnovers, TeamScore, OpponentScore, DefRushingYds, OffensivePlays)
```

```{r}
cormatrix <- rcorr(as.matrix(simplelogs))

cormatrix$r
```
Okay so I will say this, I think I just kinda proved that defense wins games by accident. But I choose Offensive plasy and Defensive plays because I wanted to see if there was a possibility that either one would correlate to the differential. There's not much of a relationship BUT it does seem like there is more of a relationship when the defense makes a turnover. There's also some kind of relationship when the offense has more penalties. It's interesting to me because I would've figured penalties and turnovers would affect both sides of the ball not just one. 

```{r}
close_games <- logs |> filter(Differential < 7)
```

I'm lost on the last one. I'm not sure how to "define" close games. Because I would define a game that is decided by 7 points or less a close game. I'm just not sure how to do that. I did look at the code block from the basketball one we did and played around with different variables but I'm still not sure. Pls help.

### Bubble Charts 

```{r}
library(tidyverse)
```

```{r}
logs <- read_csv("https://dwillis.github.io/sports-data-files/footballlogs22.csv")
```

```{r}
winlosslogs <- logs |> 
  mutate(
    wins = case_when(
      grepl("W", Outcome) ~ 1, 
      grepl("L", Outcome) ~ 0)
)
```

```{r}
winlosslogs |> 
  group_by(Team, Conference) |> 
  summarise(
    TotalPlays = sum(OffensivePlays), 
    TotalYards = sum(OffensiveYards), 
    DefensivePlays = sum(DefPlays), 
    DefensiveYards = sum(DefYards), 
    TotalWins = sum(wins)) |> 
  mutate(
    OffensiveYPP = TotalYards/TotalPlays, 
    DefensiveYPP = DefensiveYards/DefensivePlays) -> ypp
```

```{r}
ggplot() + 
  geom_point(
    data=ypp, aes(x=OffensiveYPP, y=DefensiveYPP)
    )
```


```{r}
ggplot() + 
  geom_point(
    data=ypp, 
    aes(x=OffensiveYPP, y=DefensiveYPP, size=TotalWins)
    )
```

make it transparent
```{r}
ggplot() + 
  geom_point(
    data=ypp, 
    aes(x=OffensiveYPP, y=DefensiveYPP, size=TotalWins),
    alpha = .3) + 
  scale_size(range = c(3, 8), name="Wins")
```

big ten 

```{r}
bigten <- ypp |> filter(Conference == "Big Ten Conference")

ggplot() + 
  geom_point(
    data=ypp, 
    aes(x=OffensiveYPP, y=DefensiveYPP, size=TotalWins), 
    color="grey", 
    alpha=.5) + 
  geom_point(
    data=bigten, 
    aes(x=OffensiveYPP, y=DefensiveYPP, size=TotalWins), 
    color="red")
```

```{r}
ypp |> 
  ungroup() |> 
  summarise(
    offense = mean(OffensiveYPP), 
    defense = mean(DefensiveYPP)
    )
```

```{r}
ggplot() + 
  geom_point(
    data=ypp, 
    aes(x=OffensiveYPP, y=DefensiveYPP, size=TotalWins), 
    color="grey", 
    alpha=.5) + 
  geom_point(
    data=bigten, 
    aes(x=OffensiveYPP, y=DefensiveYPP, size=TotalWins), 
    color="red") + 
  geom_vline(xintercept = 5.700174) + 
  geom_hline(yintercept = 5.55143)
```

load packages and library

```{r}
library(ggrepel)
```

```{r}
ggplot() + 
  geom_point(
    data=ypp, 
    aes(x=OffensiveYPP, y=DefensiveYPP, size=TotalWins), 
    color="grey", 
    alpha=.5) + 
  geom_point(
    data=bigten, 
    aes(x=OffensiveYPP, y=DefensiveYPP, size=TotalWins), 
    color="red") + 
  geom_vline(xintercept = 5.700174) + 
  geom_hline(yintercept = 5.55143) +
  geom_text_repel(
    data=bigten, 
    aes(x=OffensiveYPP, y=DefensiveYPP, label=Team)
  )
```

chart but with labels 
```{r}
ggplot() + 
  geom_point(
    data=ypp, 
    aes(x=OffensiveYPP, y=DefensiveYPP, size=TotalWins), 
    color="grey", 
    alpha=.5) + 
  geom_point(
    data=bigten, 
    aes(x=OffensiveYPP, y=DefensiveYPP, size=TotalWins), 
    color="red") + 
  geom_vline(xintercept = 5.700174) + 
  geom_hline(yintercept = 5.55143) + 
  geom_text_repel(
    data=bigten, 
    aes(x=OffensiveYPP, y=DefensiveYPP, label=Team)
  ) +
  labs(title="Is Maryland moving up to the Big Ten's best?", subtitle="The Terps offense and defense puts it among ranked teams in the conference.", caption="Source: NCAA | By Derek Willis")  + theme_minimal() + 
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title = element_text(size = 8), 
    plot.subtitle = element_text(size=10), 
    panel.grid.minor = element_blank()
    )
```

