```{r}
library(tidyverse)
```
```{r}
vbmatches_2023 <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensVolleyballData/main/data/ncaa_womens_volleyball_matchstats_2023.csv")
```


```{r}
goodvb <- vbmatches_2023 |> 
  select(team, kills, digs, hit_pct)
```

```{r}
glimpse(vbmatches_2023)
```
I'm not gonna lie, I don't know what is going on but I'll do my best 
```{r}

```

Before
```{r}
perset <- vbmatches_2023 |> 
  mutate(
    ptsperset = pts/s,
    hit_pctperset = hit_pct/s,
    defensive_hit_pctperset = defensive_hit_pct/s
  ) |>
  select(team, opponent, hit_pctperset, ptsperset, defensive_hit_pctperset)
```
Okay, so I added opponent because I feel like some teams from match run different defenses/offenses so depending on the night, a team's performance can be impacted. For example, Pittsburgh hit a .559 against Montana. Either Pitt was just hitting that well that day or Montana really had zero block on Pitt. 

I'm making a new block of code so you can see what I did before and after. 

After
```{r}
perset <- vbmatches_2023 |> 
  mutate(
    killsperset = kills/s,
    digsperset = digs/s,
    hit_pctperset = total_attacks/s
  ) |>
  select(team, killsperset, digsperset, hit_pctperset )
```

```{r}
persettotals <- goodvb |> 
  group_by(team) |> 
  summarise(
    KillspersetAvg = mean(kills), 
    digspersetAvg = mean(digs), 
    hit_pctpersetAvg = mean(hit_pct)
    ) 
```


Before
```{r}
teamzscore <- perset |>
  group_by(team, opponent) |> 
  mutate(
    ptszscore = as.numeric(scale(ptsperset, center = TRUE, scale = TRUE)),
    hit_pctzscore = as.numeric(scale(hit_pctperset, center = TRUE, scale = TRUE)),
    defensive_hit_pctzscore = as.numeric(scale(defensive_hit_pctperset, center = TRUE, scale = TRUE)),
    TotalZscore = ptszscore + hit_pctzscore + defensive_hit_pctzscore
  ) |> ungroup() |> na.omit(NaN)
```
I found it interesting that there was a lot of NA. So I'll take it as the metrics I used weren't as good. I think this only for this season so maybe adding last season would make the metrics a little better. I also don't know if Na is a good thing or bad thing, I'm not sure if I'm using the right metrics for a zscore at least. 


After
```{r}
teamzscore <- perset |>
  group_by(team) |> 
  mutate(
    killszscore = as.numeric(scale(killsperset, center = TRUE, scale = TRUE)),
    digszscore = as.numeric(scale(digsperset, center = TRUE, scale = TRUE)),
    hit_pctzscore = as.numeric(scale(hit_pctperset, center = TRUE, scale = TRUE)),
    TotalZscore = killszscore + digszscore + hit_pctzscore
  ) |> ungroup()
```


```{r}
head(teamzscore)
```

```{r}
teamzscore |> arrange(desc(TotalZscore))
```

```{r}
teamzscore |> 
  filter(team == "Maryland Terrapins, Terps") |> 
  arrange(desc(TotalZscore)) |>
  select(team, TotalZscore)
```
I'm having a little trouble here so I might switch the metrics back to see if maybe it's the metrics that are taking out some teams? I'm not sure why some teams are being deleted off the table/some teams aren't showing up on the team side but they are for opponent. I changed the metrics and I got the same result. 

I filtered the table for just Maryland and it didn't come up but once I did it for opponent it filtered all of Maryland's games. I'm also getting NaN again and I even changed the metrics so, I'm confused as to why that's happening. But some teams, like I did Xavier in the team filter, that will filter in team. So yeah, help pls because I feel like I messed this up. 


Update: I fixed the notebook and I didn't have NaN's anymore yay and I was able to find Maryland.