```{r}
library(tidyverse)
```

```{r}
set.seed(1234)

simulations <- rbinom(n = 1000, size = 20, prob = .317)

table(simulations)
```
n = 1000 simulations. How we read it:  how often it happens and how likely it is to occur. 

Men's basketball 
```{r}
set.seed(1234)

simulations <- rbinom(n = 1000, size = 15, prob = .326)

hist(simulations)
```
```{r}
table(simulations)
```

Lacrosse
```{r}
library(dplyr)

set.seed(1234)

simulations <- rbinom(n = 1000, size = 1, prob = .377)

four_in_a_row <- mean(simulations == 1 & lead(simulations, 1) == 1 & lead(simulations, 2) == 1 & lead(simulations, 3) == 1)

odds <- 1/four_in_a_row
```
Breanna Stewart
```{r}
set.seed(1234)

simulations <- rbinom(n = 1000, size = 17, prob = .465)

hist(simulations)
```
```{r}
table(simulations)
```

### Beeswarm plots

```{r}
library(tidyverse)
library(ggbeeswarm)
library(ggrepel)
library(ggalt)
```


```{r}
set.seed(1234)
```


```{r}
players <- read_csv("https://thescoop.org/sports-data-files/wbb_players_2023.csv")
```

```{r}
head(players)
```

```{r}
activeplayers <- players |> filter(mp>0, position != "NULL") 
```


```{r}
activeplayers <- activeplayers |> mutate(tspct=pts/(2*(fga+0.44*fta)))
```

calculate qt
```{r}
summary(activeplayers$fga)
```
182 came from 3rd QT
```{r}
shooters <- activeplayers |> filter(fga > 187)
```


```{r}
ggplot() + geom_beeswarm(data=shooters, aes(x=position, y=tspct), color="grey")
```
```{r}
umd <- activeplayers |> 
  filter(team == "Maryland") |> 
  filter(fga>187) |> 
  arrange(desc(tspct))
```

```{r}
ggplot() + 
  geom_beeswarm(
    data=shooters, 
    groupOnX=TRUE, 
    aes(x=position, y=tspct), color="grey") + 
  geom_beeswarm(
    data=umd, 
    groupOnX=TRUE, 
    aes(x=position, y=tspct), color="red")
```

```{r}
ggplot() + 
  geom_beeswarm(
    data=shooters, 
    groupOnX=TRUE, 
    aes(x=position, y=tspct), color="grey") + 
  geom_beeswarm(
    data=umd, 
    groupOnX=TRUE, 
    aes(x=position, y=tspct), color="red") +
  geom_text_repel(
    data=umd, 
    aes(x=position, y=tspct, label=player))
```

Other options

geom_quasirandom
```{r}
ggplot() + 
  geom_quasirandom(
    data=shooters, 
    groupOnX=TRUE, 
    aes(x=position, y=tspct), color="grey") + 
  geom_quasirandom(
    data=umd, 
    groupOnX=TRUE, 
    aes(x=position, y=tspct), color="red") +
  geom_text_repel(
    data=umd, 
    aes(x=position, y=tspct, label=player))
```

geom_jitter
```{r}
ggplot() + 
  geom_jitter(
    data=shooters, 
    aes(x=position, y=tspct), color="grey") + 
  geom_jitter(
    data=umd, 
    aes(x=position, y=tspct), color="red") +
  geom_text_repel(
    data=umd, 
    aes(x=position, y=tspct, label=player))
```
### Encircling points on a scatterplot

```{r}
players <- read_csv("https://raw.githubusercontent.com/dwillis/hhs-snapshots/main/data/player_totals_20231130.csv") |> filter(mp > 0)
```
Players who have played so far

```{r}
ggplot() + geom_point(data=players, aes(x=mp, y=pts))
```
```{r}
topscorers <- players |> filter(pts > 175)
```


```{r}
ggplot() + 
  geom_point(data=players, aes(x=mp, y=pts), color="grey") +
  geom_point(data=topscorers, aes(x=mp, y=pts), color="black")
```
a good way to make people pay attention to the players that are seperated from the pack. 

```{r}
ggplot() + 
  geom_point(data=players, aes(x=mp, y=pts), color="grey") + 
  geom_point(data=topscorers, aes(x=mp, y=pts), color="black") + 
   geom_text(data=topscorers, aes(x=mp, y=pts, label=full_name), hjust = 0, vjust=1) +
  geom_encircle(data=topscorers, aes(x=mp, y=pts), s_shape=0, expand=0, colour="red")
```
```{r}
ggplot() + 
  geom_point(data=players, aes(x=mp, y=pts), color="grey") + 
  geom_point(data=topscorers, aes(x=mp, y=pts), color="black") + 
  geom_encircle(data=topscorers, aes(x=mp, y=pts), s_shape=.5, expand=.03, colour="red") +
  geom_text(aes(x=275, y=275, label="Top scorers")) + 
  labs(title="Caitlin Clark Alone At Top", subtitle="The Iowa star is by far the top scorer among all NCAA players", x="Minutes", y="Points") + 
  theme_minimal() + 
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title = element_text(size = 8), 
    plot.subtitle = element_text(size=10), 
    panel.grid.minor = element_blank()
    )
```
note: to point out a specific dot, create a data frame with that person
