---
title: "Women's Soccer Data"
output: html_notebook
---

We'll use this notebook to explore some NCAA women's soccer data and [our questions](https://docs.google.com/spreadsheets/d/1kTSrnl1URmcexIWLDMWqDnp3q-c2OS45uifnhjp10n4/edit#gid=0). First, let's load the tidyverse library:

```{r}
library(tidyverse)
```

We'll want to start with 2023 match data, which is located in [this GitHub repository](https://github.com/dwillis/NCAAWomensSoccerData). We can load the match data directly from GitHub using `read_csv`:


```{r}
matches_2023 <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensSoccerData/main/data/ncaa_womens_soccer_matchstats_2023.csv")

matches_2022 <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensSoccerData/main/data/ncaa_womens_soccer_matchstats_2022.csv")
```

Let's take a look at this data to see what it contains, using `head` and `glimpse`:

```{r}
head(matches_2023)
glimpse(matches_2023)
```

Now, onto the questions using aggregates, which means `group_by` and `summarize`:

1. How many shots in total has Maryland attempted and how many have been on goal? Calculate a percentage of shots on goal. 3.61

```{r}
matches_2023|>
  group_by(team)|>
  summarise(
    total_shots = sum (sh_att),
    total_on_goal = sum(so_g),
    percent = (total_on_goal/total_shots)*100
  ) |>
  arrange(desc(percent))
```

2. What about the same shot stats from last season's first four games? You'll need to load the matches from 2022 - just change the season in the URL - and filter by date to find the first four games. I think 47??

```{r}
matches_2022|>
  filter (date < '2022-09-04')|>
  filter (team == 'Maryland Terrapins, Terps') |>
  group_by(team) |>
  summarise(
    total_shots = sum(sh_att),
    total_on_goal = sum(so_g),
    percent = (total_on_goal/total_shots)*100
  ) |>
  arrange(desc(percent))
```


3. What about draws - Maryland started 2022 with 4 ties. Have any other teams had as many this season? 1374

```{r}
matches_2022|>
  group_by(outcome)|>
  summarise(
    total = n()
  )
```


4. How many total corner kicks has Maryland had so far and how many have its opponents have? Calculate the difference between the two totals for every team. (I don't know if I did that right)

```{r}
matches_2023|>
  group_by(corners)|>
  summarise(
    total = n()
  )
```
## Significance Tests:

Hypothesis: the is more socring in women's soccer this season than last season.

Add a column to both dataframes that contains the total_goals for each match. To do this, we'll use mutate and will save the results back to the same dataframe.

```{r}
matches_2022 <- matches_2022 |> mutate (total_goals = goals + defensive_goals)

matches_2023 <- matches_2023 |> mutate (total_goals = goals + defensive_goals)

```

```{r}
t.test(matches_2022$total_goals, matches_2023$total_goals)

```
There wasn't a large difference between the two seasons. However, the hypothesis is correct that women's soccer team did score more in the current season than last. I would explain this data an audience that women's soccer are scoring more than in the previous games.

```{r}
matches_2022_early <- matches_2022 |> filter(date < '2022-09-14')
```

```{r}
t.test(matches_2022_early$total_goals, matches_2023$total_goals)
```

It's not a huge difference again but there's slight improvement of how many goals are being scored from year to year. I would explain this to an audience that this season women's soccer are taking more goals.




## Correlations

```{r}
correlations <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensSoccerData/main/data/ncaa_womens_soccer_matchstats_2023.csv")
```

```{r}
newcorrelations <- correlations |>
  mutate(
    differential = goals - defensive_goals
    )
```

```{r}
newcorrelations |> summarise(correlation = cor(differential, fouls, method="pearson"))
```


```{r}
newcorrelations |>
  summarise(correlation = cor(differential, yellow_cards, method="pearson"))
```

```{r}
fit <- lm(differential ~ fouls, data = newcorrelations)
summary(fit)
```
Fit = how much do fouls predict he outcomes in games

```{r}
newcorrelations |>
  summarise(correlation = cor(differential, so_g, method="pearson"))
```
0.68 is closer to 1, therefore a positve relationship

```{r}
net <- lm(differential ~ so_g, data = newcorrelations)
summary(net)
```

```{r}
(0.445467*11)+-2.530441
```

```{r}
newcorrelations |>
  summarise(correlation = cor(differential, sh_att, method="pearson"))
```

```{r}
net <- lm(differential ~ sh_att, data = newcorrelations)
summary(net)
```

```{r}
(0.232868*11)+-2.833383
```
Okay, I just kind of plugged in some columns that would work so I looked at shot attempts. I realized as I was doing that some teams that are more aggressive on defense tend to great more offensive plays. So, if I'm reading it correctly, Maryland shouldn't have won the Navy game based on shot attempts. I think it's interesting to look at how many attempts a team is actually taking to see if they are more likely to win games.
