## NCAA Football Attendance 

Let's load the tidyverse.

```{r}
library(tidyverse)
```

And read in the attendance data:

```{r}
attendance <- read_csv("https://dwillis.github.io/sports-data-files/attendance.csv")
```

Using the tidyverse's `pivot_longer` function, turn the currently wide data into longer data and save it as a new dataframe called `attendance_longer`:

```{r}
attendance_longer <- attendance |> pivot_longer(cols = starts_with("20"), names_to = "Year", values_to = "Attendance")
```

Let's try to figure out whether football attendance has returned to "normal" by 2022, defining "normal" as 2019 figures. For that we'll need to isolate those two years using `filter()` and get rid of any NAs:

```{r}
attendance_two_years <- attendance_longer |> filter(Year == 2019 | Year == 2022) |> na.omit()
```

Now, in order to calculate a percentage change between those two years, we'll need to go back to wider data and again ditch teams where we don't have both years:

```{r}
attendance_wide <- attendance_two_years |> pivot_wider(names_from = Year, values_from = Attendance) |> na.omit()
```

Finally, we'll add a percentage change column to that dataframe (remember, it's (New-Old)/Old) and arrange it to see the results.

```{r}
attendance_wide <- attendance_wide |> mutate(change = ((`2022`-`2019`)/`2019`)*100) |> arrange(desc(change))
```

Some of these clearly are influenced by performance on the field. Below, describe the most interesting results (positive or negative) and what you think might be going on with them. Feel free to do some googling and describe what other data would be helpful to clarify what the best stories here might be. What surprises you about these results?

**ANSWER**
I think knowing what specfically was the change in the the enviroment and why did some schools saw an increase in attendance. I was surprised to see Rutgers and UTSA in the top 10 however, I know Rutgers got a new head coach two years ago. UTSA is made up of four campuses so maybe other people started to care about football? I would like to know what makes people go to games in those places that had a huge in attendence like Rutgers, East Carolina, etc. Also, a lot of the colleges are in colleges where the weatehr is warmer so maybe being stuck in A/C all day was annoying to people who had it?? I don't know, I would like to know why people went.



