```{r}
library(tidyverse)
```

```{r}
logs <- read_csv("https://raw.githubusercontent.com/Sports-Roster-Data/womens-volleyball/main/rosters_2023_cleaned.csv")
```

```{r}
head(logs)
```

```{r}
logs |>
  filter(!is.na(state_clean)) |>
  select(team, primary_position, state_clean, name)
```

```{r}
logs |>
  group_by(state_clean) |>
  filter(!is.na(state_clean)) |>
  summarise(
    states = n()) |>
  arrange(desc(states))
```
With the new roaster data, the top ten from my first presentation has changed slightly. Wisconsin is actually now in the top 10 where I think previously they were maybe top 20. Same thing with Minn. and North Carolina. I think looking at this across all of NCAA helped in some way. Interestingly, California and Texas still lead the way by numbers. If I could find a way to break it down by division, that would be interesting. I wouldn't be surprised by the numbers if California and Texas are leading in D1. 
### Positions

```{r}
vbpositions <- logs |>
  select(primary_position, state_clean) |> 
  na.omit()
```

```{r}
vbpositions |>
  group_by(primary_position, state_clean) |>
  summarise(
    positions = n()) |> 
      arrange(desc(positions))
```
So, some of these numbers aren't shocking. Outsides are the primary position and come at an abundance no matter the state. But I did find it interesting that California produced the most aobut of Middle Blockers and Setters and Texas produces the most amount DS and Right Side. I have to wonder if that's due to the fact that they are able to play year round so players are able to develop in all areas of the game sooner. 
###Team and State

```{r}
teams <- read_csv("https://raw.githubusercontent.com/Sports-Roster-Data/womens-volleyball/main/teams.csv")
```


```{r}
logs_with_teams <- logs |> inner_join(teams, join_by("team", "ncaa_id")) 
```

P.s: If i could see the schools by division again that would be cool. 
```{r}
logs_with_teams <- logs |> inner_join(teams, join_by("team", "ncaa_id")) 
```

For Texas
```{r}
logs_with_teams |>
  filter(state_clean == 'TX') |> 
  group_by(team) |> 
  summarize(count = n()) |> 
    arrange(desc(count))
```
For CA
```{r}
logs_with_teams |>
  filter(team_state == 'CA', state_clean == 'CA', division == 'I') |> 
  group_by(team) |> 
  summarize(count = n()) |> 
    arrange(desc(count))
```
For all schools
```{r}
logs_with_teams |>
  filter(division == "III") |> 
  group_by(team, team_state, state_clean) |> 
  summarize(count = n()) |> 
    arrange(desc(count)) |> na.omit()
```
So changing the division, it looks like smaller schools recruit in state where Divison I schools tend to go out of state to look for their talent. 

```{r}
logs_with_teams |>
  filter(division == "I") |> 
  filter(conference == "Big Ten") |>
  group_by(team, team_state, state_clean) |> 
  summarize(count = n()) |> 
    arrange(desc(count)) |> na.omit()
```

```{r}
logs_with_teams |>
  filter(division == "I") |> 
  filter(conference == "SEC") |>
  group_by(team, team_state, state_clean) |> 
  summarize(count = n()) |> 
    arrange(desc(count)) |> na.omit()
```
### Looking at Top Schools in NCAA Tournment 

```{r}
logs_with_teams |>
  filter(division == "I") |> 
  group_by(team) |> 
  summarize(count = n_distinct(state_clean)) |> 
    arrange(desc(count)) |> na.omit()
```
A look at unique states each school has. (keep working with this, break down between Big conference and HBCU)
Louisville
```{r}
logs_with_teams |>
  filter(division == "I") |> 
  filter(conference == "ACC") |>
  filter(team == "Louisville") |>
  group_by(team_state, state_clean) |> 
  summarize(count = n()) |> 
    arrange(desc(count)) |> na.omit()
```
Pittsburgh
```{r}
logs_with_teams |>
  filter(division == "I") |> 
  filter(conference == "ACC") |>
  filter(team == "Pittsburgh") |>
  group_by(team_state, state_clean) |> 
  summarize(count = n()) |> 
    arrange(desc(count)) |> na.omit()
```

Chart (come back to this)
```{r}
all_states <- logs_with_teams |>
  group_by(state_clean) |> 
  summarize(count = n()) |> 
    arrange(desc(count))

ggplot() +
  geom_histogram(data=all_states, aes(y=count)) +
  facet_wrap(~state_clean)
  
```


### Per Capita 

```{r}
states <- read_csv("http://goodcsv.com/wp-content/uploads/2020/08/us-states-territories.csv")
```

```{r}
usstates <- states |>
  filter(!is.na(`Population (2019)`)) |>
  select(state) |>
  top_n(50)
```
I'm playing around with this CSV file with the states. I'm working on making sure it's just US states and not territories. The population data is from 2019, so if I can find a more updated one, I will. 
